Update and evaluation of the ozone dry deposition in the Oslo~CTM3 v1.0

### 2019-04-05 version_2/response_version_1.tex

\item {\color{blue}  P22 Table 3:}
  \begin{itemize}
  \item {\color{blue}  Why the deposition values for ocean + ice + land do to add up to the total
    values reported, for all simulations?}
    Are we right to assume that the referee is trying to state that the values for ocean, ice, and land do
    not add up to the total global ozone dry deposition? If that is the case, he is right, for we exclusively
    selected gridboxes associated to more then 98\% with the three surface types (ocean, ice, and land), while
    \emph{total} comprises all gridboxes. We admit that this is not clear.
    We shall clarify this in the table caption and the revised manuscript, repeat the computation of the values
    with the proper weighting, and adapt the numbers.
    We may also add that \emph{ice} in this analyis only
    referes to regions at high latitudes that are permanently covered by ice and snow and hence does not take
    sea ice and mountain glaciers into account. In the Oslo~CTM3, as written elsewhere in the manuscript, we
    compute dry deposition on ice and snow based on meteorological data. Therefore, the given values in Table~3
    comprise ozone dry deposition on sea ice in the case of \emph{ocean} and on snow covered land during winter
    in case of \emph{land}.

  \item {\color{blue}  The new land-based deposition values are much lower than what has
    been reported in previous studies (e.g. Hardacre et al., 2015) and the authors largely
    attribute this to the changes in the updated scheme for the desert surface type. However,
    the paper does not provide any observational support to back this up. }
    \begin{itemize}
    \item {\color{blue}  Are there any relevant deposition measurements (velocity or flux) that can be used for this purpose?}
      The only paper we know of is a study by G\"{u}sten~et~al.~(1996) in which measurements of ozone concentrations
      and fluxes onto the Sahara desert are described and dry deposition velocities deduced. We can include a more
      thorough discussion of our results with respect to the observed fluxes by G\"{u}sten~et~al.~(1996) in the
      revision of our manuscript. 
    \item {\color{blue}  At least, some comparison with ozone measurements (or even $\mathrm{O_3}$ reanalyses)
      should be provided for this surface type (and perhaps others) to see if the model is
      heading in the right direction with the updated deposition scheme.}
      Thank you for the advise. We will look into this.
    \end{itemize}
  \item {\color{blue}  It will also be useful to report the global ozone burden from the various
    simulations.}
    Since Table~3 is already at maximum width with respect to the the page width, we shall show the global
    tropospheric ozone burden in a seperate table (\ref{tab:trop_ozone_burden}). If we compare our results
    with Stevenson~et~al.~(2006)
    ($344\pm 39\,\mathrm{Tg}$) and the number given in IPPC AR5 (2013) ($337\pm 23\,\mathrm{Tg}$), we find
    that the ozone burden in the Oslo~CTM3 is higher then the model average from the start (\emph{Wesely scheme}).
    The implementation of the \emph{EMEP scheme}
    increases the tropospheric burden by roughly 8\,\% (compare \emph{Wesely type} and EMEP\_full).
    \begin{table*}[h]
        \caption{Annual mean tropospheric ozone burden for all experiments and $1 \sigma$ standart deviation.}
        \centering
        \begin{tabular}{lrcl}
          \hline
          Experiment & \multicolumn{3}{c}{Trop. $\mathrm{O_3}$}\\
          &  \multicolumn{3}{c}{(Tg)}\\
          \hline
          Wesely type & 361 & $\pm$ & 21\\
          EMEP\_full & 392 & $\pm$ & 28\\
          EMEP\_offLight & 388 & $\pm$ & 26\\
          EMEP\_offPhen & 392 & $\pm$ & 27\\
          EMEP\_SWVL4 & 402 & $\pm$ & 31\\
          EMEP\_ppgs & 392 & $\pm$ & 28\\
          EMEP\_ppgssh & 391 & $\pm$ & 27\\
          EMEP\_ppgssh\_ice & 403 & $\pm$ & 31\\
          EMEP\_ppgs\_2005 & 386 & $\pm$ & 26\\
          \hline
        \end{tabular}
        \label{tab:trop_ozone_burden}
    \end{table*}
  \end{itemize}

### 2019-04-05 DryDep.tex
Continue to write the response. Added Torp ozone burden table.

Finally, we take a look at the different global as well as hemispheric dry deposition sinks for ozone (Table~\ref{tab:ozone_sinks}). Despite its vastness, the ocean amounts only to about $33\,\unit{\%}$ of the global ozone sink due to dry deposition in the Oslo~CTM3 with operative EMEP scheme (\emph{EMEP\_full}), while permanently ice and snow covered regions account for $1.5\,\unit{\%}$. The remainder is deposited to land surfaces of which deserts might be the most neglected in process-modeling. The total annual dry deposition in the EMEP scheme is about one-third below the multi-model-mean result by \citet{ACP:Hardacre2015}. But also the results of \citet{ACP:Luhar2017, ACP:Luhar2018} yield a $(19-27)\,\unit{\%}$ lower ozone dry deposition than the models participating in the model intercomparison, with deposition to ocean ranging between $(12-21)\,\unit{\%}$ of the total annual ozone dry deposition. In particular, \citet{ACP:Luhar2018} found that current model estimates of ozone dry deposition to the ocean may be three times too high compared to their analyisis. This implies that the ozone dry deposition to the ocean the Oslo~CTM3 is too high as well.
%
\begin{table*}[t]
\caption{Total ozone dry deposition for the respective model experiment in \unit{Tg\,a^{-1}}. The global ozone dry deposition has been weighted by ocean, ice and, land fraction in each gridbox, respectively. \emph{Ice} herein refers to regions at high latitudes that are permanently covered by ice and snow.}
\begin{tabular}{lccccccccc|cccr}
\tophline
\multirow{3}{*}{Simulation} & \multicolumn{3}{c}{Ocean} & \multicolumn{3}{c}{Ice} & \multicolumn{3}{c}{Land} & \multicolumn{3}{c}{Total} & $\Delta^\dagger$\\
& NH & SH & Global & NH & SH & Global & NH & SH & Global & NH & SH & Global\\
& \multicolumn{3}{c}{(\unit{Tg\,a^{-1}})} & \multicolumn{3}{c}{(\unit{Tg\,a^{-1}})} & \multicolumn{3}{c}{(\unit{Tg\,a^{-1}})} & \multicolumn{3}{c}{(\unit{Tg\,a^{-1}})} & (\unit{\%})\\
\middlehline
Wesely\_type & 159.7 & 147.8 & 307.5 & 3.6 & 6.3 & 9.9 & 446.6 & 193.7 & 640.3 & 612.4 & 347.9 & 960.2 & 46.8\\
EMEP\_full & 107.6 & 105.2 & 212.8 & 2.5 & 5.5 & 8.0 & 296.6 & 135.0 & 431.6 & 408.6 & 245.6 & 654.2 & 0.0\\
EMEP\_offLight & 110.0 & 105.8 & 215.8 & 2.5 & 5.4 & 7.9 & 337.1 & 156.0 & 493.0 & 451.3 & 267.2 & 718.6 & 9.8\\
EMEP\_offPhen & 108.0 & 105.3 & 213.2 & 2.5 & 5.5 & 8.0 & 301.6 & 139.1 & 440.7 & 413.9 & 249.9 & 663.7 & 1.5\\
EMEP\_SWVL4 & 109.2 & 107.0 & 216.2 & 2.6 & 5.5 & 8.1 & 303.6 & 138.2 & 441.8 & 417.2 & 250.7 & 667.9 & 2.1\\
EMEP\_ppgs & 107.6 & 105.2 & 212.7 & 2.5 & 5.5 & 8.0 & 299.5 & 135.3 & 434.9 & 411.4 & 246.0 & 657.4 & 0.5\\
EMEP\_ppgssh & 108.3 & 105.4 & 213.6 & 2.5 & 5.5 & 8.0 & 306.8 & 142.8 & 449.6 & 419.4 & 253.6 & 673.0 & 2.9\\
EMEP\_ppgssh\_ice & 107.4 & 105.3 & 212.7 & 1.2 & 1.6 & 2.8 & 303.0 & 144.1 & 447.2 & 413.0 & 251.1 & 664.1 & 1.5\\
EMEP\_ppgs\_2005 & 106.9 & 103.5 & 210.4 & 2.6 & 5.4 & 8.0 & 297.9 & 131.8 & 429.7 & 409.2 & 240.7 & 649.9 & -0.7\\
\bottomhline
\end{tabular}
\belowtable{$^\dagger$: Difference of global annual total relative to \emph{EMEP\_full}.}% Table Footnotes
\label{tab:ozone_sinks}
\end{table*}
%

>>>>

Finally, we take a look at the different global as well as hemispheric dry deposition sinks for ozone (Table~\ref{tab:ozone_sinks}). Despite its vastness, the ocean amounts only to about $33\,\unit{\%}$ of the global ozone sink due to dry deposition in the Oslo~CTM3 with operative EMEP scheme (\emph{EMEP\_full}), while permanently ice and snow covered regions account for $1.5\,\unit{\%}$. The remainder is deposited to land surfaces of which deserts might be the most neglected in process-modeling. The total annual dry deposition in the EMEP scheme is about one-third below the multi-model-mean result by \citet{ACP:Hardacre2015}. But also the results of \citet{ACP:Luhar2017, ACP:Luhar2018} yield a $(19-27)\,\unit{\%}$ lower ozone dry deposition than the models participating in the model intercomparison, with deposition to ocean ranging between $(12-21)\,\unit{\%}$ of the total annual ozone dry deposition. In particular, \citet{ACP:Luhar2018} found that current model estimates of ozone dry deposition to the ocean may be three times too high compared to their analyisis. This implies that the ozone dry deposition to the ocean the Oslo~CTM3 is too high as well.
%
\begin{table*}[t]
  \caption{Total ozone dry deposition for the respective model experiment in \unit{Tg\,a^{-1}}. The global ozone dry deposition has been weighted by ocean, ice and, land fraction in each gridbox, respectively. \emph{Ice} herein refers to regions at high latitudes that are permanently covered by ice and snow.}
  \begin{tabular}{lccccccccc|cccr}
    \tophline
    \multirow{3}{*}{Experiment} & \multicolumn{3}{c}{Ocean} & \multicolumn{3}{c}{Ice} & \multicolumn{3}{c}{Land} & \multicolumn{3}{c}{Total} & $\Delta^\dagger$\\
    & NH & SH & Global & NH & SH & Global & NH & SH & Global & NH & SH & Global\\
    & \multicolumn{3}{c}{(\unit{Tg\,a^{-1}})} & \multicolumn{3}{c}{(\unit{Tg\,a^{-1}})} & \multicolumn{3}{c}{(\unit{Tg\,a^{-1}})} & \multicolumn{3}{c}{(\unit{Tg\,a^{-1}})} & (\unit{\%})\\
    \middlehline
    Wesely\_type & 159.7 & 147.8 & 307.5 & 3.6 & 6.3 & 9.9 & 446.6 & 193.7 & 640.3 & 612.4 & 347.9 & 960.2 & 46.8\\
    EMEP\_full & 107.6 & 105.2 & 212.8 & 2.5 & 5.5 & 8.0 & 296.6 & 135.0 & 431.6 & 408.6 & 245.6 & 654.2 & 0.0\\
    EMEP\_offLight & 110.0 & 105.8 & 215.8 & 2.5 & 5.4 & 7.9 & 337.1 & 156.0 & 493.0 & 451.3 & 267.2 & 718.6 & 9.8\\
    EMEP\_offPhen & 108.0 & 105.3 & 213.2 & 2.5 & 5.5 & 8.0 & 301.6 & 139.1 & 440.7 & 413.9 & 249.9 & 663.7 & 1.5\\
    EMEP\_SWVL4 & 109.2 & 107.0 & 216.2 & 2.6 & 5.5 & 8.1 & 303.6 & 138.2 & 441.8 & 417.2 & 250.7 & 667.9 & 2.1\\
    EMEP\_ppgs & 107.6 & 105.2 & 212.7 & 2.5 & 5.5 & 8.0 & 299.5 & 135.3 & 434.9 & 411.4 & 246.0 & 657.4 & 0.5\\
    EMEP\_ppgssh & 108.3 & 105.4 & 213.6 & 2.5 & 5.5 & 8.0 & 306.8 & 142.8 & 449.6 & 419.4 & 253.6 & 673.0 & 2.9\\
    EMEP\_ppgssh\_ice & 107.4 & 105.3 & 212.7 & 1.2 & 1.6 & 2.8 & 303.0 & 144.1 & 447.2 & 413.0 & 251.1 & 664.1 & 1.5\\
    EMEP\_ppgs\_2005 & 106.9 & 103.5 & 210.4 & 2.6 & 5.4 & 8.0 & 297.9 & 131.8 & 429.7 & 409.2 & 240.7 & 649.9 & -0.7\\
    \bottomhline
  \end{tabular}
  \belowtable{$^\dagger$: Difference of global annual total relative to \emph{EMEP\_full}.}% Table Footnotes
  \label{tab:ozone_sinks}
\end{table*}
%
